---
title: "Homework 3"
author: "Teshawna Badu"
date: 10/20/2021
output: github_document
---

I'm an R Markdown document!

```{r setup}
library(tidyverse)
library(dplyr)
library(p8105.datasets)
library(janitor)
library(readr)
```

## Question 1 
```{r}
data("instacart")
```

Description: 
Instacart is an American company that operates in grocery delivery and pick up services. The `instacart` dataset contains data from 2017. It has `r nrow(instacart)` observations and `r ncol(instacart)` variables. The total number of data points in this dataset is `r nrow(instacart)*ncol(instacart)`. The key variables in this dataset include `r colnames(instacart)`. 

### How many aisles are there and which are the most ordered from?
```{r, message = FALSE, warning = FALSE}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```
There are 134 `aisles`. The aisles with the most ordered-from items are 
`fresh vegetables`, `fresh fruits`, and followed by `packaged vegetables fruits`. 

### make a plot 
This plot shows the number of items ordered per aisle, limited to 10,000 items ordered.
```{r aisles_plot}
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  ggplot(aes(forcats::fct_reorder(aisle, (n)), n)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Number of items ordered per aisle",
    x = "Aisle name",
    y = "Number of items ordered",
    caption = "Data from instacart limited to aisles with more than 10,000 items ordered"
  )
```

### make a table 
This table shows three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”, Including the number of times each item is ordered. 

```{r aisles_table}
instacart %>%
  filter(aisle == "baking ingredients" | 
           aisle == "dog food care" | 
           aisle == "packaged vegetables fruits") %>%
 count(aisle, product_name) %>%
  arrange(desc(n)) %>%
  group_by(aisle) %>%
  slice(1:3) %>%
  arrange(desc(n)) %>%
  group_by(aisle) %>%
  knitr::kable(caption = "Top  3  Items Per Aisle")
```

The table shows that `organic baby spinach`, `organic raspberries`, and `organic blueberries` 
are the top three ordered items from the `packaged vegetables fruits` aisle. 
From the `baking ingredients` aisle, `light brown sugar`, `pure baking soda`, and `cane sugar` are the top three ordered items. The top three ordered items from the `dog food care` aisle are`snack sticks chicken & rice recipe dog treats`, `organix chicken & brown rice recipe`, and `small dog biscuits`.


### 2x7 table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r apples_ice_cream_table}
 instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
  ) %>%
  rename( "Sunday" = "0","Monday" = "1", "Tuesday" = "2", "Wednesday" = "3", "Thursday" = "4", "Friday" = "5", "Saturday" = "6") %>% 
  knitr::kable(caption = "Mean Hour of Day Ordered")
```
`Coffee Ice Cream` is ordered, on average, later in the day than `Pink Lady Apples`.


## Question 2
I will clean the data `brfss_smart2010`, formatting to use appropriate variable names, 
focusing on the `Overall Health` topic, and only including factored responses 
ordered from`Poor` to `Excellent`.

```{r clean_brfss}
data("brfss_smart2010")
brfss_cleaned_version = brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic %in% c("Overall Health")) %>%
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))) %>%
  arrange(response) %>%
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%
  separate(locationdesc, into = c('state', 'location'), sep = ' - ')
```


### In 2002, which states were observed at 7 or more locations?
```{r states2002}
states_2002 = brfss_cleaned_version %>% 
  filter(year == "2002") %>% 
  group_by(state) %>% 
  distinct(location) %>% 
  count(state) %>% 
  filter(n >= 7) %>% 
  select(state)
```

### In 2010, which states were observed at 7 or more locations?
```{r states2010}
states_2010 = brfss_cleaned_version %>% 
  filter(year == "2010") %>% 
  group_by(state) %>% 
  distinct(location) %>% 
  count(state) %>% 
  filter(n >= 7) %>% 
  select(state)
```

The states that were observed at 7 or more locations in 2002 were `r states_2002$state`. 
The states that were observed at 7 or more locations in 2010 were `r states_2010$state`.

### Plot 
I will construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state.Additionally I will make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r brfss_excellent}
Excellent_df = brfss_cleaned_version %>%
  filter(response %in% "Excellent") %>%
  select(year, locationabbr, data_value) %>%
  unique() %>%
  na.omit() %>%
  group_by(locationabbr, year) %>%
  mutate(
    avg_data_value = mean(data_value)) %>%
  select(year, locationabbr, avg_data_value) %>%
  distinct()
Excellent_df %>%
  ggplot(aes(x = year, y = avg_data_value)) +
  geom_line(aes(group = locationabbr, color = locationabbr)) +
  labs(
    title = "Average data value over time for statess",
    x = "Year",
    y = "Average Data Value") +
  guides(col = guide_legend("State")) +
  theme(legend.position = "right")
```

### Make two panel showing 
I will make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r panels}
brfss_panels = brfss_cleaned_version %>% 
  filter(year %in% c(2006,2010),
  state == "NY") 
ggplot(brfss_panels, aes(x = response, y = data_value, fill = response)) + 
  geom_boxplot() + 
  facet_grid(. ~ year) +
  labs(
    title = "Distribution of data_value for responses in NY",
    x = "Response",
    y = "data_value",
    caption = "Data from the brfss dataset"
  )
```

## Question 3 
```{r}
accel_data = read_csv("./accel_data.csv")
```

I will load, tidy, and otherwise wrangle the data. The final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r}
accel_df = accel_data %>% 
  pivot_longer(
    cols = activity.1:activity.1440,
    names_to = "activity_number",
    values_to = "activity_counts",
    names_prefix = "activity.",
  ) %>% 
  mutate(
    is_weekend = (day == "Saturday" | day == "Sunday")
    )
```

The tidied dataset `accel_df` contains `r nrow(accel_df)` rows (or observations) and `r ncol(accel_df)` variables. The dataset's `r ncol(accel_df)` variables include the following: 
`r colnames(accel_df)`

### Inspecting daily activity 
Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
total_activity = 
  accel_df %>% 
  group_by(day_id) %>% 
  summarize(total_activity_counts = sum(activity_counts))
knitr::kable(total_activity)
```

The total activity counts are does not have a distinct pattern, the counts fluctuate every couple of days. 


### Panel 
Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r fig.width = 8, fig.height = 4}
accel_df %>% 
  mutate(activity_number = as.numeric(activity_number)) %>% 
  group_by(day, activity_number) %>% 
  summarize(avg_value = mean(activity_counts)) %>% 
  ggplot(aes(x = activity_number, y = avg_value, color = day)) +
  geom_smooth(se = FALSE) +
  scale_x_discrete(limit = c(360,720,1080,1440), 
                   labels = c( "6", "12", "18", "24")) +
  labs(
    title = "Average 24 Hour Activity for Days of the Week",
    x = "Activity Number (hrs)",
    y = "Average Activity Counts",
    color = "Day"
  )
```

On average, activity counts for all days in a week are at lowest during 23:50 p.m. to 6:00 a.m. There are peaks of average activity counts at around 10:30 a.m of Thurdsay and around 21:00 p.m of Friday. The low levels may be due to sleeping and inability to be active, while the high peaks may be due to daytime and having some free time to exercise. 
